Clustering and Parameter Estimation for the Popularity Adjusted Block Model

### 2020-07-14 summary.Rmd
cleaned up some typos and such

---
title: "Clustering and Parameter Estimation for the Popularity Adjusted Block Model"
output:
  pdf_document:
    citation_package: natbib
    number_sections: yes
# output: html_document
# geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,vlined,ruled]{algorithm2e} 
- \setcitestyle{numbers,square}
bibliography: misc.bib
abstract: |
  In this paper, we connect two probabolistic models for graphs, the Popularity 
  Adjusted Block Model (PABM) and the Generalized Random Dot Product Graph 
  (GRDPG) and use properties established in this connection to aid in 
  clustering and parameter estimation. In particular, we note that the PABM 
  can be represented as latent positions such that points within the same 
  cluster lie on a subspace, and the subspaces that represent each 
  cluster are orthogonal to one another. Using this property as well as the 
  asymptotic properties of Adjacency Spectral Embedding (ASE) of the GRDPG, we 
  are able to establish theoretical asymptotic results of our clustering and 
  parameter estimation methods for the PABM. 
---

@@

One popular probabolistic model for graphs with underlying communities is the 
Stochastic Block Model (SBM), which assumes that the edge probability from 
one cluster to another follows a Beronulli distribution with a fixed 
probability for each pair of communities. 
The Popularity Adjusted Block Model (PABM) was introduced by 
\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the SBM to 
address the heterogeneity of edge probabilities within and between communities 
or clusters while still maintaining the community structure.

The Random Dot Product Graph (RDPG) \cite{athreya2017statistical} is another 
probabolistic model with Beronulli edge probabilities. Under this model, each 
vertex of the graph can be represented by a point in latent space such that 
the edge probability between any pair of vertices is given by their 
corresponding dot product in the latent space. The SBM is equivalent to a 
special case of the RDPG model in which all vertices in a given community share 
the same position in the latent space. An analogous property exists for the 
PABM not under the RDPG model but under the *Generalized* Random Dot Product 
Graph model.

@@

The adjacency spectral embeddings of both $A$ and $P$ are not unique. In 
particular, let $Z$ be the ASE of $P$. Then $P = Z I_{3, 1} Z^\top$. However, 
for any $Q \in \mathbb{O}(3, 1)$, 
$(ZQ) I_{3, 1} (ZQ)^\top = Z (Q I_{3, 1} Q^\top) Z^\top = Z I_{3, 1} Z^\top = P$,
so $ZQ$ is also a valid ASE Of $P$. If we can find $Q \in \mathbb{O}(3, 1)$ 
such that $ZQ = XU$, we can compute the parameters $\{\lambda^{(kl)}\}$ 
directly. Furthermore, if we instead use the ASE of the adjacency matrix $A$, 
$\exists Q \in \mathbb{O}(3, 1)$ such that $\max\limits_i ||Q \hat{z}_i - XU||$ 
is minimized (and goes to zero under a probabolistic model for
$\{\lambda^{(kl)}\}$'s).

>>>>

---
title: "Clustering and Parameter Estimation for the Popularity Adjusted Block Model"
output:
  pdf_document:
    citation_package: natbib
    number_sections: yes
# output: html_document
# geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,vlined,ruled]{algorithm2e} 
- \setcitestyle{numbers,square}
bibliography: misc.bib
abstract: |
  In this paper, we connect two probabilistic models for graphs, the Popularity 
  Adjusted Block Model (PABM) and the Generalized Random Dot Product Graph 
  (GRDPG) and use properties established in this connection to aid in 
  clustering and parameter estimation. In particular, we note that the PABM 
  can be represented as latent positions such that points within the same 
  cluster lie on a subspace, and the subspaces that represent each 
  cluster are orthogonal to one another. Using this property as well as the 
  asymptotic properties of Adjacency Spectral Embedding (ASE) of the GRDPG, we 
  are able to establish theoretical asymptotic results of our clustering and 
  parameter estimation methods for the PABM. 
---

@@

One popular probabilistic model for graphs with underlying communities is the 
Stochastic Block Model (SBM), which assumes that the edge probability from 
one cluster to another follows a Beronulli distribution with a fixed 
probability for each pair of communities. 
The Popularity Adjusted Block Model (PABM) was introduced by 
\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the SBM to 
address the heterogeneity of edge probabilities within and between communities 
or clusters while still maintaining the community structure.

The Random Dot Product Graph (RDPG) \cite{athreya2017statistical} is another 
probabilistic model with Beronulli edge probabilities. Under this model, each 
vertex of the graph can be represented by a point in latent space such that 
the edge probability between any pair of vertices is given by their 
corresponding dot product in the latent space. The SBM is equivalent to a 
special case of the RDPG model in which all vertices in a given community share 
the same position in the latent space. It has also been shown that similar 
probabilistic graph models, (such as the Mixed Membership Stochastic Block Model
\cite{rubindelanchy2017statistical}). An analogous property exists for the 
PABM not under the RDPG model but under the *Generalized* Random Dot Product 
Graph model.

@@

The adjacency spectral embeddings of both $A$ and $P$ are not unique. In 
particular, let $Z$ be the ASE of $P$. Then $P = Z I_{3, 1} Z^\top$. However, 
for any $Q \in \mathbb{O}(3, 1)$, 
$(ZQ) I_{3, 1} (ZQ)^\top = Z (Q I_{3, 1} Q^\top) Z^\top = Z I_{3, 1} Z^\top = P$,
so $ZQ$ is also a valid ASE Of $P$. If we can find $Q \in \mathbb{O}(3, 1)$ 
such that $ZQ = XU$, we can compute the parameters $\{\lambda^{(kl)}\}$ 
directly. Furthermore, if we instead use the ASE of the adjacency matrix $A$, 
$\exists Q \in \mathbb{O}(3, 1)$ such that $\max\limits_i ||Q \hat{z}_i - XU||$ 
is minimized (and goes to zero under a probabilistic model for
$\{\lambda^{(kl)}\}$'s).
